---
title: "Getting started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(tictoc)
mustashe::clear_stash()
```

The purpose of the 'mustashe' R package is to save objects that result from some computation, then load the object from file the next time the computation is performed.
In other words, the first time a chunk of code is evaluated, the output can be *stashed* for the next time the code chunk is run.

## Basic example

Below is a brief example outlining the use of the primary function from the package, `stash()`.
First we must load the 'mustashe' library.

```{r}
library(mustashe)
```

Say we are performing a long-running computation (simulated here using `Sys.sleep()` to pause for a few seconds) that produces the object named `x`.
The name of the object to stash `"x"` and the code itself are passed to `stash()` as follows.
(I used the package ['tictoc'](https://CRAN.R-project.org/package=tictoc) to time the execution of the code.)

```{r}
tic("long-running computation")

stash("x", {
  Sys.sleep(5)
  x <- 5
})

toc()
```

'mustashe' tells us that the object was stashed, and we can see that `x` was successfully assigned the value 5.

```{r}
x
```

Say we are done for the day, so we close RStudio and go home.
When we return the next day and continue on the same analysis, we really don't want to have to run the same computation again since it will have the same result as yesterday.
Thanks to 'mustashe', the code is not evaluated and, instead, the object `x` is loaded from file.

```{r}
tic("long-running computation")

stash("x", {
  Sys.sleep(5)
  x <- 5
})

toc()
```

That's the basic use case of 'mustashe'!
Any issues and feedback can be submitted [here](https://github.com/jhrcook/mustashe/issues).
Continue reading below for explanations of other useful features of 'mustashe'.

### Why not use 'ProjectTemplate's `cache()` function?

Originally I tried to use the `cache()` function from ['ProjectTemplate'](https://CRAN.R-project.org/package=ProjectTemplate), but ran into a few problems.

The first was, to use it without modification, I would need to be using the 'ProjectTemplate' system for my whole analysis project. It first checks if all of the expected directories and components are in place, throwing an error when they are not.

```r
ProjectTemplate::cache("x")

#> Current Directory: mustashe is not a valid ProjectTemplate directory because one or more mandatory directories are missing.  If you believe you are in a ProjectTemplate directory and seeing this message in error, try running migrate.project().  migrate.project() will ensure the ProjectTemplate structure is consistent with your version of ProjectTemplate.
#> Change to a valid ProjectTemplate directory and run cache() again.

#> Error in .quietstop():
```

I then tried copying the source code for the `cache()` function to my project and tweaking it to work (mainly removing internal checks for 'ProjectTemplate' system).
I did this and thought it was working: on the first pass it would cache the result, and on the second it would load from the cache.
However, in a new session of R, it would not just load from the cache, but, instead, evaluate the code and cache the results.
After a bit of exploring the `cache()` source code, I realized the problem was that 'ProjectTemplate' compares the current value of the object to be cached with the object that is cached.
Of course, this requires the object to be in the environment already, which it is in a 'ProjectTemplate' system after running `load.project()` because that loads the cache (lazily) into the R environment.
I do not want this behavior, and thus the caching system used by 'ProjectTemplate' was insufficient for my needs.

That said, I *heavily* relied upon the code for `cache()` when creating `stash()`.
This would have been far more difficult to do without reference to 'ProjectTemplate'.


## Features

There are two major features of the `stash()` function from 'mustashe' not covered in the basic example above:

1. 'mustashe' "remembers" the code passed to `stash()` and will re-evaluate the code if it has changed.
2. Dependencies can be explicitly linked to the stashed object so that the code is re-evaluated if the dependencies change.

These two features are demonstrated below.

### 'mustashe' "remembers" the code

If the code that creates an object changes, then the object itself is likely to have changed.
Thus, 'mustashe' "remembers" the code and re-evaluates the code if it has been changed.
Here is an example, again using 'tictoc' to indicate when the code is evaluated.

```{r}
tic()
stash("a", {
  Sys.sleep(3)
  a <- runif(5)
})
toc()
```

```{r}
tic()
stash("a", {
  Sys.sleep(3)
  a <- runif(10)
})
toc()
```

**However, 'mustashe' is insensitive to changes in comments and other style-based adjustments to the code.**
In the next example, a comment has been added, but we see that the object is loaded from the stash.

```{r}
stash("a", {
  Sys.sleep(3)
  # Here is a new comment.
  a <- runif(10)
})
```

And below is the code from a horrible person, but 'mustashe' still loads the object from the stash.

```{r}
stash("a", {
  Sys.sleep(3)

  # Here is a comment.

  a <- runif(10) # Another comment
})
```

### Dependencies

Dependencies can be explicitly linked to an object to make sure that if they change, the stashed object is re-evaluated.
"Dependency" in this case could refer to data frames that are used to create another (e.g. summarizing a data frame's columns), inputs to a function, etc.

The following demonstrates this with a simple example where `x` is used to calculate `y`.
By passing `"x"` to the `depends_on` argument, when the value of `x` is changed, the code to create `y` is re-evaluated

```{r}
x <- 1

stash("y", depends_on = "x", {
  y <- x + 1
})

# Value of `y`
y
```

The second time this is run without changing `x`, the value for `y` is loaded from the stash.

```{r}
stash("y", depends_on = "x", {
  y <- x + 1
})
```

However, if we change the value of `x`, then the code is re-evaluated and the stash for `y` is updated.

```{r}
x <- 100

stash("y", depends_on = "x", {
  y <- x + 1
})

# Value of `y`
y
```

Multiple dependencies can be passed as a vector to `depends_on`.

```{r}
stash("y", depends_on = c("x", "a"), {
  y <- x + a
})
```


## Other API features

### Unstashing and clearing stash

To round up the explanation of the 'mustashe' package, the stash can be cleared using `unstash()` and specific stashes can be removed using `unstash()`.

```{r}
unstash("a")
```

```{r}
clear_stash()
```

### Functional interface

In the examples above, `stash()` does not return a value (actually, it invisibly returns `NULL`), instead assigning the result of the computation to an object named using the `var` argument.
Frequently, though, a return value is desired.
This behavior can be induced by setting the argument `functional = TRUE`.

```{r}
b <- stash("b", functional = FALSE, {
  rnorm(5, 0, 1)
})
b
```

```{r}
b <- stash("b", functional = TRUE, {
  rnorm(5, 0, 1)
})
b
```

### Functions as dependencies

The `stash()` function can take other functions as dependencies.
The body and formals components of the function object are checked to see if they have changed.
(More information on the structure of function objects in R can be found in Hadley Wickham's [*Advanced R* - Functions: Function components](http://adv-r.had.co.nz/Functions.html#function-components).)

As an example, suppose you have a script with the following code.
It is run, and the value of 5 is stashed for `a` and it is dependent on the function `add_x()`.

```{r}
add_x <- function(y, x = 2) {
  y + x
}

stash("a", depends_on = "add_x", {
  a <- add_x(3)
})
a
```

You continue working and change the function `add_x()` to use the default value of 5 instead of 2.
This change will cause the code for `a` to be re-run and `a` will be assigned the value 8.
Note that the code in the `code` argument for `stash()` did not change, the code was re-run because a dependency changed.

```{r}
add_x <- function(y, x = 5) {
  y + x
}

stash("a", depends_on = "add_x", {
  a <- add_x(3)
})
a
```

### Using `stash()` in functions

Because of the careful management of R environments, `stash()` can be used inside of functions.
In the example below, note that the stashed object will depend on the value of the `magic_number` object *in the function*.

```{r}
magic_number <- 10
do_data_science <- function() {
  magic_number <- 5
  stash("rand_num", depends_on = c("magic_number"), {
    runif(1, 0, 10)
  })
  return(rand_num)
}

do_data_science()
```

Changing the value of the `magic_number` object in the global environment will not invalidate the stash.

```{r}
magic_number <- 11
do_data_science()
```

### Stashing results of sourcing a R script

It is also possible to stash the results of sourcing and R script.
The contents of the script are an implicit dependency for the stash, so if the script changes, it will be re-sourced the next time around.
It is also possible to include additional dependencies using the `depends_on`
parameter in the same way as with a regular stash.
 
The natural behavior of the `source()` function is maintained by returning the last evaluated value in the script.

```{r}
# Write a temporary R script.
temp_script <- tempfile()
write("print('Script to get 5 letters'); sample(letters, 5)", temp_script)

x <- stash_script(temp_script)
x
```

```{r}
x2 <- stash_script(temp_script)
x2
```


## Configuration

### Using ['here'](https://here.r-lib.org) to create file paths

The ['here'](https://here.r-lib.org) package is useful for handling file paths in R projects, particularly when using an RStudio project.
The main function, `here::here()`, can be used to create the file path for stashing an object by setting the 'mustashe' configuration option with the `config_mustashe()` function.

```{r}
config_mustashe(use_here = TRUE)
```

This behavior can be turned off, too.

```{r}
config_mustashe(use_here = FALSE)
```

### Other options

Defaults for the `verbose` and `functional` (see above) arguments of stashing functions can also be configured.
For example, you can have the functions run silently and return the result by default.

```{r}
config_mustashe(verbose = FALSE, functional = TRUE)
```

---

### Contact

Any issues and feedback  on 'mustashe' can be submitted [here](https://github.com/jhrcook/mustashe/issues).
Alternatively, I can be reached through the contact form on my [website](https://joshuacook.netlify.app) or on Twitter [\@JoshDoesa](https://twitter.com/JoshDoesa)

```{r, echo=FALSE}
unlink(".mustashe", recursive = TRUE)
```
